### 2015-12-10 prb_resub/referee_response/response.txt
rev-10/11/2015

Dear Editors,

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE FIRST REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------
	This paper deals with the study of entropies in classical spin models.
	The focus is on the Sherrington-Kirkpatrick model, a simple model that
	exhibits spin glass physics. Both analytical and numerical results are
	presented, and compared.

	Overall I found the paper interesting, and clearly written. Such
	classical entropies had only been studied in nearest neighbor spin
	models, so this is a useful addition to the existing literature.

RESPONSE:
========
	We thank the referee for considering our work interesting and 
	clearly written.

REFEREE:
-------

	There is one issue that needs to be addressed I think. In part II the
	classical Renyi entropy is defined by Eq. (5), and this is what is
	studied in the rest of the paper. It is not clear to me which is the
	underlying probability distribution, i.e it is not clear to me that
	this equals Eq. (2), because of the long-range interactions. This
	point needs to be clarified, as using the term Renyi entropy would be
	a slight abuse of terminology, and the reader should be made aware of
	that.

RESPONSE:
========

	We apologise for not being very clear about the equivalence between (5) 
	and (2). Eq. (5) is equivalent to Eq. (2) even with the long range 
	interactions. This is now clarified in the first paragraph in sec II. 
	Notice that the only physical effect of long range interactions is that 
	there is no area law in contrast with local models, as it is discussed 
	extensively in the manuscript.

REFEREE:
-------

	I also have a couple of minor remarks and suggestions, which I list
	below:

	1) In the long introduction, several values of \beta are mentioned. It
	might be helpful to recall the values of $\beta_c(w)$.

RESPONSE:
========

	In order to meet the suggestion of the second referee we moved the summary of 
	results in the conclusions. Therefore, in the new version of the manuscript 
	the referee suggestion is implemented in the conclusions. First, we should 
	clarify that $\beta$ always refers to the explicit temperature in a single 
	replica, and never to the *effective* temperature on the booklet. We thank 
	the referee for noticing that in the introduction we do not mention  explicitly 
	the values of the critical temperature for the model on the booklet. We now 
	mention explicitly that the critical temperature is a decreasing function of 
	the booklet aspect ratio $\omega$ and that $\beta_c=1,1/2$ for $\omega=0,1$. 
	Notice that that all the  values of beta that we mention in the summary of 
	the results (\beta=1,\beta=3) correspond to the low-temperature phase of the 
	model on the booklet.

REFEREE:
-------

	2) Maybe at the $\forall i \in A$ also in equation (8) ?

RESPONSE:
========

	We thank the referee for noticing that. We moved the $\forall i...$  
	in equation (8).

3) Before (40), the sentence "..., one obtains a standard Landau
theory" reads awkwardly.

RESPONSE:
========

	We removed the sentence in question.

REFEREE:
-------
	4) Could the authors comment somewhere on the Monte Carlo statistical
	errors when they are not shown?
	This is: 
	-Fig 8 (hidden behind some of the points, so you could say smaller than 
	point size in caption).
	-Fig 9 (same as 8)
	In the other figures the errors bars are fairly clear, but Fig 5 might be 
	non-obvious at first glance

RESPONSE:
========

	We thank the referee for the useful comment. In the captions of Fig 5, Fig 8, 
	and Fig 9 we now mention explicitly that the error bars are smaller than the 
	symbol size.

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE SECOND REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------

	In the manuscript BG13275, the authors study the Reny entropy and
	mutual information in the Sherrington-Kirkpatrick model. Using the
	replica method, the authors constructed the replica-symmetric solution
	and 1-step replica-symmetry-breaking solutions. They are solved
	numerically and are compared with the results of the Monte Carlo
	simulations. It was shown that the one-step RSB solution is enough to
	find a good agreement with the Monte Carlo result. The aspect-ratio
	and temperature dependences of the entropy and mutual information are
	analyzed in detail to see the scaling in terms of the system size and
	the critical properties.

RESPONSE:
========

	We thank the referee for the accurate summary of our results.

REFEREE:
-------

	The manuscript is written very clearly and the analytical calculations
	as well as numerical ones look reliable. The authors followed the
	standard procedure of the replica analysis and several ansatz such as
	(28) look reasonable.

RESPONSE:
========

	We thank the referee for finding our paper well written and our analysis 
	reliable.

REFEREE:
-------

	Although the analysis looks quite reasonable, the authors do not
	discuss the results thoroughly nor their physical meaning. Therefore,
	I can recommend the publication if the authors clarify following
	several questions.

	(1). What is the significance to treat infinite-range models in the
	problem of mutual information? As the authors mentioned in the
	manuscript, there is no notion of space and locality in infinite-range
	models. It is not clear for me to consider the booklet geometry in
	such systems. It would be helpful if the authors answer this simple
	question in the manuscript.

RESPONSE:
========

	We thank the referee for the opportunity to clarify this point. The 
	definition of the mutual information is based on formula (2) in the 
	manuscript. Formula (2) is the standard definition of the classical 
	Renyi entropy, which in the limit $n\to 1$ become the Shannon entropy. 
	Formula (2) and the Shannon entropy can be used in systems that do not 
	have the notion of space and locality. The two subsystems A and B 
	appearing in (2) can be interpreted as two complementary groups of 
	spins, without any need for an underlying spatial bipartition. Thus, 
	the quantities that we address in our study, e.g., the mutual information, 
	are well-grounded from an information theoretical point of view. This 
	is now stressed in the introduction below formula (2) with the sentence

		" For local spin models the bipartition correspond to a spatial 
		separation between the spins. However, the definition~\eqref{renyi-intro} 
		can be used  in models with no notion of space, where it quantifies 
		the correlation between two groups of spins rathen than two spatial 
		regions."

	Moreover, (2) can be interpreted as ratios of partition functions defined 
	on the booklet. This is an exact correspondence, and it is not related to 
	any notion of locality. To clarify this point we now show explicitly the 
	equivalence between the booklet construction and (2) in the beginning of 
	section 2. From the derivation it is clear that the booklet construction 
	is meaningful also in mean-field models.

	Besides its meaning from the information theory point of view, from the 
	physics perspective the mutual information provides a lot of information 
	about critical systems without disorder. One of the most striking example is 
	provided by the results in Ref. [29] in the manuscript. We believe that it is 
	a natural and interesting question to study the mutual information in 
	disordered systems, especially in presence of glassy physics.

	(i) For instance, an interesting physics question per se it is to understand 
	what are the fingerprints, if any, of glassy physics in the mutual 
	information or in information-based quantities in general.

	(ii) A related question is whether 
	information motivated quantities could provide alternative good criticality 
	witnesses for glassy transitions. To this purpose we should stress that 
	even in the simple S-K while it is relatively easy to determine numerically 
	the critical point of the paramagnetic-glassy transition using standard 
	tools from critical phenomena, such as the Binder cumulant. In presence 
	of a transverse magnetic field the transition persists but the critical 
	point changes as a function of the field strength (Almeida-Thouless line). 
	In presence of a field it is however much more demanding, as it is shown 
	numerically for instance in

		M. Picco, F. Ritort, and M. Sales, Eur. Phys. J. B 19, 565 (2001)

	Even more difficult it is to clarify whether there is an Almeida-Thouless 
	line in short-range disordered spin models.

	Our work, although for the S-K model, provides the first results in this 
	direction. Concerning (i) for the S-K model a related question is about 
	the effect of the replica symmetry breaking on information motivated 
	quantities. We clarified that while in the low-temperature phase of the 
	model the replica symmetry breaking is reflected in the mutual information.

	As correctly observed by the referee the 1RSB approximation is enough to 
	obtain good agreement with the Monte Carlo simulations. However, we should 
	stress that this is not true at arbitrarily low temperatures, as 
	at some point the effects of the higher order replica symmetry breaking 
	should become visible.

	Concerning (ii)  our work shows that for mean-field spin glasses the 
	mutual information does not show any crossing for different system sizes 
	at the critical point. This implies that the muutual information as defined 
	in the manuscript cannot be used as a good indicator of glassy phase 
	transitions.

	One interesting question would be whether this "failure" is due to the 
	non locality of the model or to the presence of disorder. This could be 
	investigated in disordered models exhibiting a glassy transition. This 
	can be done only numerically as there is no exactly solvable model with 
	spin glass transition and local interactions. Moreover, one should consider 
	a three dimensional model. In fact, to our knowledge there are no glassy 
	transitions at finite temperature in disordered spin models. This certainly 
	goes beyond the scope of the present work.

	Another possibility, as observed by the referee, it is that one has to 
	focus on the subleading corrections to the mutual information in order 
	to obtain a clear signature of the glassy transition. In general, subleading 
	corrections to the entropies have been investigated a lot in the literature 
	in the context of topolocical phases of matter, systems breaking continuous 
	symmetries and bipartitions with sharp corners.

	Figure 9 in our work provides some hints in this direction. Clearly the scaling 
	corrections are dramatically different in the paramagnetic and in the glassy 
	phase, decaying much more slowly in the latter. We regret we did not stress 
	enough this point in the manuscript. We now added the sentence in the introduction

		"We numerically observede that while in the paramagnetic phase the 
		scaling correction to the muatual information decay as 1/L, a much 
		slower decay as 1/L^2/3 is observed in the glassy phase."

	The study of these scaling corrections, although possible using the same 
	techniques employed for the regular S-K model, it is demanding as it requires 
	using the corrections to the saddle point treatment presented in section II.

REFEREE:
-------

	Is there any other work treating the booklet geometry in mean field 
	systems? For example, what is obtained for the simple, pure Ising model 
	with infinite-range interaction?

RESPONSE:
========

	The only study we are aware of is in

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

	In this work the authors investigate the mutual information in the Lipkin-Meshkov-Glick 
	(LMG) model. In particular, they provide results for the case with zero magnetic field, 
	where the model becomes equivalent to a classical infinite-range Ising model. However, 
	the authors investigate the standard mutual information which is obtained from the 
	full reduced density matrix of the subsystem A.

	In contrast with our results the authors find that the mutual information is finite 
	away from the critical point and diverges logarithmically at the critical point.

	In our case this logarithmic correction might be subleading to the volume law.

	TO BE COMPLETED

	About the maximum

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P010011 (2011)

REFEREE:
-------
	(2). The definition of the mutual information in the present case is
	not clear for me. Please describe how to define and calculate S(B) and
	S(A or B) in the SK model.

RESPONSE:
========

	We thank the referee for this useful comment. 
	Eq. (2), and later Eq. (5) via the replica trick, are both quite explicit. 
	Combined with Eq. (4) this precisely described the mutual information in 
	the SK model. The non-locality of the interactions does 
	not affect the definition compared with the usual one for local models. 
	The definition of region A is however less obvious, and in the case of 
	this work, we simply take it as an arbitrary set of points, rather than 
	one that is spatially connected. As it is clear from the manuscript 
	the best way to calculate the mutual information is using (4) calculating 
	the three different contributions S(A) S(B) and S(AUB) separately. These 
	correspond to ratios of partition functions on the booklet with different 
	aspect ratio omega. As this was not clearly stated in the manuscript we now 
	have written it explicitly in section 2

		"The mutual information ${\mathcal I}_n(A,B,\beta)$ is obtained in terms of the booklet 
		partition functions using Eq.~\eqref{renyi} and Eq.~\eqref{MI}. 
		%%%
		Clearly, $S_n(B)$ (cf. Eq.~\eqref{MI}) is obtained from~\eqref{renyi} and exchanging 
		$A$ and $B$. Notice also that in~\eqref{renyi}  $Z(A\cup B,n,\beta)\equiv Z(n\beta)$ 
		is the partition function of the model on the booklet with all the $n$ sheets identified, 
		equivalently on a single sheet but at temperature $n\beta$."

	From the numerical point of view, the strategy to evaluate the ratios of partition 
	functions in (5) is outlined Appendix B.

REFEREE:
-------

	(3). The authors summarized that the entropy is proportional to the
	system size by using the numerical analysis. Since the entropy is
	calculated from the logarithm of the partition function, the result
	looks trivial for me. As far as I understand, the subleading term is
	important to find the boundary law behavior in the local (non
	mean-field) models. (The leading term is canceled out by the
	definition of the mutual information). Do the authors consider that
	the subleading term is not important in the present system? I expect
	that some universal properties exist in the subleading terms.

RESPONSE:
========

	We agree that in general the leading term is proportional to the 
	system size is an obvious result. In the manuscript we report 
	the entropy only as a sanity check of our calculations.  We also 
	agree that it is likerly that the subleading terms contain 
	more universal information. We thank the referee for raising this 
	point. Although the calculation of the subleading corrections to 
	the entropy and the mutual information is possible in principle, 
	it requires going beyond the saddle point approximation described 
	in section III C. This should be straightforward in the paramagnetic 
	phase although it would require a considerable amount of work in 
	below the critical point. Notice that the same problem is 
	challenging in the regular SK model. We think that this goes 
	beyond the scope of the paper and it is suitable for future work.

	PUT REFERENCE TO PAPERS STUDYING FINITE SIZE IN SK

REFEREE:
-------

	(4). The authors study the aspect-ratio dependence of the result
	closely. It is clearly understood that the limits omega=0 and 1
	reduces to the standard results. This means that we can find a result
	interpolating their limits in the present problem. I do not expect any
	interesting behavior between the limits. What did the authors expect
	as the result and what do the authors consider the obtained results
	interesting, other than the crossing behavior?

RESPONSE:
========

	We agree that omega=0 and 1 are trivial results, at least in terms of 
	the standard SK model, and that one would expect that the results between 
	these two points should vary smoothly, and this is reflected in the data. 
	What was not obvious was the form of the interpolation between these two 
	points. The same statement of the referee would apply to local models.  
	However, as it is clear from Ref [29], in that case the shape function of the mutual 
	information contains universal behavior if the parent model is critical and 
	described by a conformal field theory. In our study we investigate the 
	shape function in a non-trivial disordered model. Notice that this functions 
	have been also investigated in clean (withoud disorder) mean field models 
	in 
		O. Cohen, V. Rittenberg, T. Sadhu J Phys A 48 055002 (2015)

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

	In that respect, 
	we believe it was worth examining the aspect ratio behavior in the SK model. 
	The difference of course is that since the gluing procedure is a bulk effect, 
	the entire system may go through a phase transition as a function of omega, 
	rather than being able to view the system as two parts in contact with each 
	other where one part is critical. In this light, it might be some non-constant 
	temperature cut in (omega,beta) that shows the most interesting behavior.

REFEREE:
-------

	(5). This comment is related to the previous one (4). Strictly
	speaking, I do not see any surprising behavior in the result. What is
	the significance of the present study? Did the authors find anything
	on the information properties of the SK model?

RESPONSE:
========

	The non surprising result would have been if 1-RSB would have reproduced the results 
	of the Monte Carlo simulation. The fact that 2-RSB was necessary shows the non-trivial 
	nature of the transition in the replicated geometry, and represents an expanded set of 
	assumptions compared to the base model. The procedure for handling disorder, usually done 
	via integration over fictitious replicas, in the context of the replica trick, where one 
	has physical replicas, is non-trivial, but grounding such theoretical calculations with 
	numerics is necessary. In this respect, we see the SK model with the replica trick as a 
	base case for studying replica models  in glassy systems, in the same way that the SK 
	model itself is a prototype for understanding standard glassy systems.

REFEREE:
-------

	(6). The authors only focused on the result at n=2. It would be
	interesting if the authors mention the n-dependence of the result.
	Especially, the limit n=1 would give a nontrivial result and is
	interesting. I expect that it wouldn't be difficult to analyze the
	replica symmetric solution.

RESPONSE:
========

	TO BE DONE.

REFEREE:
-------

	(7). Although the manuscript is written very clearly, it is somewhat
	too long. For example, the authors summarized the results in the
	introduction, and they are described again in the final section. It is
	not necessary to summarize the results in the introduction.

RESPONSE:
========

	We would like to thank the referee for this useful comment. We now 
	summarize the main results of the paper in the conclusions.

REFEREE:
-------
	(8). J_{ij} should be inserted in Eq.(10).

RESPONSE:
========

	We thank the referee for spotting this. We restored the missing J_{ij} in 
	Eq. (10).

REFEREE:
-------

	(9). I am not sure whether this manuscript is suitable to PRB. I would
	think that PRE is more appropriate.

RESPONSE:
========

	Argue that similar paper appeared in PRB.

In conclusion, the manuscript will be taken into consideration for
publication if the authors clarify the physical significance of the
analysis. And, I recommend the authors to reconsider the journals to
submit the manuscript.