### 2015-12-16 prb_resub/referee_response/response.txt
16/12/2015

Dear Editors,

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE FIRST REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------
	This paper deals with the study of entropies in classical spin models.
	The focus is on the Sherrington-Kirkpatrick model, a simple model that
	exhibits spin glass physics. Both analytical and numerical results are
	presented, and compared.

	Overall I found the paper interesting, and clearly written. Such
	classical entropies had only been studied in nearest neighbor spin
	models, so this is a useful addition to the existing literature.

RESPONSE:
========
	We thank the referee for considering our work interesting and 
	clearly written.

REFEREE:
-------

	There is one issue that needs to be addressed I think. In part II the
	classical Renyi entropy is defined by Eq. (5), and this is what is
	studied in the rest of the paper. It is not clear to me which is the
	underlying probability distribution, i.e it is not clear to me that
	this equals Eq. (2), because of the long-range interactions. This
	point needs to be clarified, as using the term Renyi entropy would be
	a slight abuse of terminology, and the reader should be made aware of
	that.

RESPONSE:
========

	We apologise for not being very clear about the equivalence between (5) 
	and (2). Eq. (5) is equivalent to Eq. (2) even with the long range 
	interactions. This is now clarified in the first paragraph in sec II. 
	Notice that the only physical effect of long range interactions is that 
	there is no area law in contrast with local models, as it is discussed 
	extensively in the manuscript.

REFEREE:
-------

	I also have a couple of minor remarks and suggestions, which I list
	below:

	1) In the long introduction, several values of \beta are mentioned. It
	might be helpful to recall the values of $\beta_c(w)$.

RESPONSE:
========

	In order to meet the suggestion of the second referee we moved the summary of 
	results in the conclusions. Therefore, in the new version of the manuscript 
	the referee suggestion is implemented in the conclusions. First, we should 
	clarify that $\beta$ always refers to the explicit temperature in a single 
	replica, and never to the *effective* temperature on the booklet. We thank 
	the referee for noticing that in the introduction we do not mention  explicitly 
	the values of the critical temperature for the model on the booklet. We now 
	mention explicitly that the critical temperature is a decreasing function of 
	the booklet aspect ratio $\omega$ and that $\beta_c=1,1/2$ for $\omega=0,1$. 
	Notice that that all the  values of beta that we mention in the summary of 
	the results (\beta=1,\beta=3) correspond to the low-temperature phase of the 
	model on the booklet.

REFEREE:
-------

	2) Maybe at the $\forall i \in A$ also in equation (8) ?

RESPONSE:
========

	We thank the referee for noticing that. We moved the $\forall i...$  
	in equation (8).

REFEREE:
-------

	3) Before (40), the sentence "..., one obtains a standard Landau
	theory" reads awkwardly.

RESPONSE:
========

	We removed the sentence in question.

REFEREE:
-------
	4) Could the authors comment somewhere on the Monte Carlo statistical
	errors when they are not shown?
	This is: 
	-Fig 8 (hidden behind some of the points, so you could say smaller than 
	point size in caption).
	-Fig 9 (same as 8)
	In the other figures the errors bars are fairly clear, but Fig 5 might be 
	non-obvious at first glance

RESPONSE:
========

	We thank the referee for the useful comment. In the captions of Fig 5, Fig 8, 
	and Fig 9 we now mention explicitly that the error bars are smaller than the 
	symbol size.

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE SECOND REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------

	In the manuscript BG13275, the authors study the Reny entropy and
	mutual information in the Sherrington-Kirkpatrick model. Using the
	replica method, the authors constructed the replica-symmetric solution
	and 1-step replica-symmetry-breaking solutions. They are solved
	numerically and are compared with the results of the Monte Carlo
	simulations. It was shown that the one-step RSB solution is enough to
	find a good agreement with the Monte Carlo result. The aspect-ratio
	and temperature dependences of the entropy and mutual information are
	analyzed in detail to see the scaling in terms of the system size and
	the critical properties.

RESPONSE:
========

	We thank the referee for the accurate summary of our results.

REFEREE:
-------

	The manuscript is written very clearly and the analytical calculations
	as well as numerical ones look reliable. The authors followed the
	standard procedure of the replica analysis and several ansatz such as
	(28) look reasonable.

RESPONSE:
========

	We thank the referee for finding our paper well written and our analysis 
	reliable.

REFEREE:
-------

	Although the analysis looks quite reasonable, the authors do not
	discuss the results thoroughly nor their physical meaning. Therefore,
	I can recommend the publication if the authors clarify following
	several questions.

	(1). What is the significance to treat infinite-range models in the
	problem of mutual information? As the authors mentioned in the
	manuscript, there is no notion of space and locality in infinite-range
	models. It is not clear for me to consider the booklet geometry in
	such systems. It would be helpful if the authors answer this simple
	question in the manuscript.

RESPONSE:
========

	We thank the referee for the opportunity to clarify this point. The 
	definition of the mutual information is based on formula (2) in the 
	manuscript. Formula (2) is the standard definition of the classical 
	Renyi entropy, which in the limit $n\to 1$ become the Shannon entropy. 
	Formula (2) and the Shannon entropy can be used in systems that do not 
	have the notion of space and locality. The two subsystems A and B 
	appearing in (2) can be interpreted as two complementary groups of 
	spins, without any need for an underlying spatial bipartition. Thus, 
	the quantities that we address in our study, e.g., the mutual information, 
	are well-grounded from an information theoretical point of view. This 
	is now stressed in the introduction below formula (2) with the sentence

		" For local spin models the bipartition correspond to a spatial 
		separation between the spins. However, the definition~\eqref{renyi-intro} 
		can be used  in models with no notion of space, where it quantifies 
		the correlation between two groups of spins rathen than two spatial 
		regions."

	Moreover, (2) can be interpreted as ratios of partition functions defined 
	on the booklet. This is an exact correspondence, and it is not related to 
	any notion of locality. To clarify this point we now demonstrate explicitly 
	the equivalence between the booklet construction and (2) in the beginning of 
	section 2. From the derivation it is clear that the booklet construction 
	is meaningful also in mean-field models.

	A further direct check is provided in appendix B where we discuss (following 
	the referee suggestion) the classical Renyi entropy and the mutual information 
	for the infinite-range clean (i.e., without disorder) Ising model. For instance, 
	in Figure 12 and Figure 13 the continuous lines are obtained using the booklet 
	construction, while the symbols (rhombi) are the results using the definition 
	of the classical Renyi entropy and mutual information (formulas (2) and (4) 
	in the manuscript). The agreement, apart for finite-size corrections, 
	confirms the applicability of the booklet construction for infinite-range 
	models.

	Besides its meaning from the information theory point of view, from the 
	physics perspective the mutual information provides a lot of information 
	about critical systems without disorder. One of the most striking example is 
	provided by the results in Ref. [29] in the manuscript. We believe that it is 
	a natural and interesting question to study the mutual information in 
	disordered systems, especially in presence of glassy physics.

	(i) For instance, an interesting physics question per se it is to understand 
	what are the fingerprints, if any, of glassy physics in the mutual 
	information or in information-based quantities in general.

	(ii) A related question is whether 
	information motivated quantities could provide alternative good criticality 
	witnesses for glassy transitions. To this purpose we should stress that 
	even in the simple S-K while it is relatively easy to determine numerically 
	the critical point of the paramagnetic-glassy transition using standard 
	tools from critical phenomena, such as the Binder cumulant. In presence 
	of a transverse magnetic field the transition persists but the critical 
	point changes as a function of the field strength (Almeida-Thouless line). 
	In presence of a field it is however much more demanding, as it is shown 
	numerically for instance in

		M. Picco, F. Ritort, and M. Sales, Eur. Phys. J. B 19, 565 (2001)

	Even more difficult it is to clarify whether there is an Almeida-Thouless 
	line in short-range disordered spin models.

	Our work, although for the S-K model, provides the first results in this 
	direction. Concerning (i) for the S-K model a related question is about 
	the effect of the replica symmetry breaking on information motivated 
	quantities. We clarified that while in the low-temperature phase of the 
	model the replica symmetry breaking is reflected in the mutual information.

	As correctly observed by the referee the 1RSB approximation is enough to 
	obtain good agreement with the Monte Carlo simulations. However, we should 
	stress that this is not true at arbitrarily low temperatures, as 
	at some point the effects of the higher order replica symmetry breaking 
	should become visible.

	Concerning (ii)  our work shows that for mean-field spin glasses the 
	mutual information does not show any crossing for different system sizes 
	at the critical point. This implies that the muutual information as defined 
	in the manuscript cannot be used as a good indicator of glassy phase 
	transitions.

	One interesting question would be whether this "failure" is due to the 
	non locality of the model or to the presence of disorder. This could be 
	investigated in disordered models exhibiting a glassy transition. This 
	can be done only numerically as there is no exactly solvable model with 
	spin glass transition and local interactions. Moreover, one should consider 
	a three dimensional model. In fact, to our knowledge there are no glassy 
	transitions at finite temperature in disordered spin models. This certainly 
	goes beyond the scope of the present work.

	Another possibility, as observed by the referee, it is that one has to 
	focus on the subleading corrections to the mutual information in order 
	to obtain a clear signature of the glassy transition. In general, subleading 
	corrections to the entropies have been investigated a lot in the literature 
	in the context of topolocical phases of matter, systems breaking continuous 
	symmetries and bipartitions with sharp corners.

	Figure 9 in our work provides some hints in this direction. Clearly the scaling 
	corrections are dramatically different in the paramagnetic and in the glassy 
	phase, decaying much more slowly in the latter. We regret we did not stress 
	enough this point in the manuscript. We now added the sentence in the introduction

		"We numerically observede that while in the paramagnetic phase the 
		scaling correction to the muatual information decay as 1/L, a much 
		slower decay as 1/L^2/3 is observed in the glassy phase."

	The study of these scaling corrections, although possible using the same 
	techniques employed for the regular S-K model, it is demanding as it requires 
	using the corrections to the saddle point treatment presented in section II.

REFEREE:
-------

	Is there any other work treating the booklet geometry in mean field 
	systems? For example, what is obtained for the simple, pure Ising model 
	with infinite-range interaction?

RESPONSE:
========

	We would like to thank the referee for this very useful comment. The only study we 
	are aware of is in

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

	In this work the authors investigate the mutual information in the Lipkin-Meshkov-Glick 
	(LMG) model. In particular, they provide results for the case with zero magnetic field, 
	where the model becomes equivalent to a classical infinite-range Ising model. However, 
	the authors investigate the standard mutual information which is obtained from the 
	Shannon entropy, instead of the classical Renyi one.

	In contrast with our results the authors find that the Shannon mutual information is 
	finite at any temperature except at the critical point where it diverges logarithmically.

	Motivated by the referee observation and the results of Wilms et al we now discuss 
	in the new version of the manuscript the Renyi entropy and mutual information in the 
	clean infinite-range Ising model (in appendix B). We find that, in constrast with the 
	Shannon mutual information, the Renyi one exhibits a volume law for any temperature, 
	similar to our result for the S-K model. Moreover, I_2 exhibits intriguing subleading 
	(i.e., subextensive) contributions. Specifically, we numerically observe that these 
	corrections are logarithmically divergent with system size N. Moreover, they are 
	present at three values of the inverse temperature. Their precise behaviors is given 
	as

		beta=1/2     ---->  1/4 log(N)

		beta=1/(1+w) ---->  1/2 log(N)

		beta=1	     ---->  1/2 log(N)

	Notice that beta=1/(1+w) corresponds to the critical temperature on the 2-sheets 
	booklet with aspect ratio w.

	These new results raise the question whether similar subleading corrections are 
	present in the S-K model, as also noticed by the referee. Importantly, it would be 
	interesting to determine the prefactor of these logarithmic corrections to check 
	whether they are affected by the disorder. However, one should notice that to 
	investigate analytically this issue one should go beyond the saddle point 
	approximation, which at least in the low-temperature regioon requires knowing the 
	full replica-symmetry-broken solution of the model.

	Another possibility is to focus at beta=1/2 where, in analogy with the clean case, 
	a logarithmic correction is expected and where the high-temperature replica symmetric 
	result fully captures the saddle point behavior. The strategy would be to consider the 
	deviations from the saddle point in the Monte Carlo data. However, we checked that 
	our numerical results are not precise enough to determine the corrections and further 
	numerical effort would be needed to clarify this point.

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P010011 (2011)

REFEREE:
-------
	(2). The definition of the mutual information in the present case is
	not clear for me. Please describe how to define and calculate S(B) and
	S(A or B) in the SK model.

RESPONSE:
========

	We thank the referee for this useful comment. Eq. (2), and later Eq. (5) via the replica 
	trick, are both quite explicit. Combined with Eq. (4) they give the mutual information in 
	the SK model. The non-locality of the interactions does not affect the definition compared 
	with the usual one for local models. The definition of region A is however less obvious, 
	and in the case of this work, we simply take it as an arbitrary set of points, rather than 
	one that is spatially connected. As it is clear from the manuscript the best way to calculate 
	the mutual information is using (4) calculating the three different contributions S(A) S(B) 
	and S(AUB) separately. These correspond to ratios of partition functions on the booklet with 
	different aspect ratio w. In particular, S(AUB) is obtained from the ratio Z(w=1)/Z(w=0). 
	As this was not clearly stated in the manuscript we now have written it explicitly in 
	section 2

		"The mutual information ${\mathcal I}_n(A,B,\beta)$ is obtained in terms of the booklet 
		partition functions using Eq.~\eqref{renyi} and Eq.~\eqref{MI}. 
		%%%
		Clearly, $S_n(B)$ (cf. Eq.~\eqref{MI}) is obtained from~\eqref{renyi} and exchanging 
		$A$ and $B$. Notice also that in~\eqref{renyi}  $Z(A\cup B,n,\beta)\equiv Z(n\beta)$ 
		is the partition function of the model on the booklet with all the $n$ sheets identified, 
		equivalently on a single sheet but at temperature $n\beta$."

	From the numerical point of view, the strategy to evaluate the ratios of partition 
	functions in (5) is outlined Appendix B.

REFEREE:
-------

	(3). The authors summarized that the entropy is proportional to the system size by using the 
	numerical analysis. Since the entropy is calculated from the logarithm of the partition function, 
	the result looks trivial for me. As far as I understand, the subleading term is	important to find 
	the boundary law behavior in the local (non mean-field) models. (The leading term is canceled out 
	by the definition of the mutual information). Do the authors consider that the subleading term is 
	not important in the present system? I expect that some universal properties exist in the subleading 
	terms.

RESPONSE:
========

	We agree that in general the leading term is proportional to the system size is an obvious result. 
	In the manuscript we report the entropy only as a sanity check of our calculations.  We also agree 
	that it is likely that the subleading terms contain more universal information. We thank the referee 
	for raising this point. Although the analytical calculation of the subleading corrections to the entropy 
	and the mutual information is possible in principle, it requires going beyond the saddle point approximation 
	described in section III C. This should be straightforward in the paramagnetic phase although it would 
	require a considerable amount of work below the critical point. Notice that the same problem of 
	calculating saddle point corrections is challenging in the regular SK model. We think that this goes 
	beyond the scope of the paper and it is suitable for future work.

REFEREE:
-------

	(4). The authors study the aspect-ratio dependence of the result closely. It is clearly understood that 
	the limits omega=0 and 1 reduces to the standard results. This means that we can find a result interpolating 
	their limits in the present problem. I do not expect any interesting behavior between the limits. What 
	did the authors expect as the result and what do the authors consider the obtained results interesting, 
	other than the crossing behavior?

RESPONSE:
========

	We agree that omega=0 and 1 are trivial results, at least in the standard SK model, and that one would 
	expect that the results between these two points should vary smoothly, and this is reflected in our data. 
	However, different aspect ratios omega are needed in the booklet constructions in order to construct the 
	mutual information. We should also stress that the functional dependence of the mutual information on 
	omega is not trivial. The same statement of the referee would apply to local models. However, as it is clear 
	from Ref [29], the shape function of the mutual information contains universal behavior if the 
	parent model is critical and described by a conformal field theory. In that respect, we believe it was 
	worth examining the aspect ratio behavior in the SK model. Notice that this functions have been also 
	investigated in clean (withoud disorder) mean field models in

		O. Cohen, V. Rittenberg, T. Sadhu J Phys A 48 055002 (2015)

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

REFEREE:
-------

	(5). This comment is related to the previous one (4). Strictly
	speaking, I do not see any surprising behavior in the result. What is
	the significance of the present study? Did the authors find anything
	on the information properties of the SK model?

RESPONSE:
========

	From our answer to point 1) it should be clear now that our work is well grounded from the information 
	point of view as we are studying the standard classical Renyi entropies and mutual information. These 
	quantities have been shown to contain universal information for instance in local critical models described 
	a conformal field theory (Ref. [29]). Again, we should stress that their behavior in disordered and glassy systems has 
	not been investigated in the previous literature. We chose to focus on the SK model which is one of the few 
	exactly solved models for spin glasses and the base case for understanding glassy behavior, and its solution, 
	although the model is mean field, exhibits quite complex features.

	We have found that the Renyi entropy S_2 and the mutual information I_2 have a volume law and we provided analytical 
	results using the saddle point approximation. This can be consired as a zeroth order starting point because, as 
	correctly observed by the referee and as clear from our results on the clean infinite range Ising model (appendix B), 
	further interesting information could be contained in the finite-size effects. However, already the saddle point 
	results for S_2 and I_2 are non trivial since they reflect the replica symmetry breaking pattern of the solution 
	of the SK model. We should stress that we only provided the 1-RSB approximation for these quantities. The fact that 
	the 2-RSB is necessary at low temperatures confirms the fact that our results are interesting.

	Our results are essential in order to go beyond the saddle point, studying subextensive correctins in S_2 and I_2. 
	This is confirmed by the results that we provide in appendix B where we show that already for the clean infinite 
	range Ising model the mutual information contains logarithmically divergent (with system size) terms which reflect 
	critical behaviors. Our work raises the important question as to whether such subextensive contributions are present 
	at glassy transitions. Moreover, a related interesting question is if disorder renormalizes the prefactor of this 
	logarithmic behavior. Additionally, for the SK model it is interesting to understand whether the replica-symmetry-breaking 
	mechanism affects this prefactor. Besides being interesting in their own, our results are necessary in order to 
	understand the subleading  of S_2 and I_2.

REFEREE:
-------

	(6). The authors only focused on the result at n=2. It would be	interesting if the authors mention the n-dependence of 
	the result. Especially, the limit n=1 would give a nontrivial result and is interesting. I expect that it wouldn't 
	be difficult to analyze the replica symmetric solution.

RESPONSE:
========

	We thank the referee for asking about this point. In fact we thought about investigating the analytic continuation 
	n->1 to obtain the Shannon Renyi entropy and mutual information. The referee question is interesting especially 
	especially in order to compare with the results for the clean infinite range Ising model obtained in Wilms et al. 
	It has been shown in fact in ... that the Shannon mutual information is O(1) (i.e., not extensive) at any temperature 
	except for the critical point at beta=1 whete it diverges logarithmically with system size.

	Following the referee suggestion we now provide results for the analytic continuation n->1. First, one should observe 
	that in the limit of infinite temperature one has from formula () in our paper that

	   I_n(beta)/N = beta^2 omega(1-omega)n/2    for beta --> 0

	which implies that

	   I_1/N = beta^2 omega(1-omega)/2

	i.e., the Shannon entropy obeys a volume law in the paramagnetic phase. This is in contrast with the result in the 
	clean Ising reported in Wilms et al and it is apparently an effect of the disorder. This suggests that the subextensive 
	behavior for I_1 found for the clean system is an effect of fine tuned cancellations that do not happen in presence of 
	disorder.

	In the paper we now included the saddle point equation in the high-temperature phase (equation (31)).

REFEREE:
-------

	(7). Although the manuscript is written very clearly, it is somewhat
	too long. For example, the authors summarized the results in the
	introduction, and they are described again in the final section. It is
	not necessary to summarize the results in the introduction.

RESPONSE:
========

	We would like to thank the referee for this useful comment. We now 
	summarize the main results of the paper in the conclusions and not in the 
	introduction, while in the introduction we stress better the physical motivation 
	and the results of the work.

REFEREE:
-------
	(8). J_{ij} should be inserted in Eq.(10).

RESPONSE:
========

	We thank the referee for spotting this. We restored the missing J_{ij} in 
	Eq. (10).

REFEREE:
-------

	(9). I am not sure whether this manuscript is suitable to PRB. I would
	think that PRE is more appropriate.

RESPONSE:
========

	Argue that similar paper appeared in PRB.

In conclusion, the manuscript will be taken into consideration for
publication if the authors clarify the physical significance of the
analysis. And, I recommend the authors to reconsider the journals to
submit the manuscript.