### 2015-12-20 prb_resub/referee_response/response_v1.txt

Dear Editors,

We are submitting a revised version of the manuscript BG13275
"The classical mutual information in mean-field spin glass models".

We would like to thank the editors for their work and the referees for 
their useful comments and suggestions.

We remind that the first referee considered our work interesting and clearly 
written and recommended it for publication in its present form. The second 
referee cosidered our results reliable and the paper very clearly written.
However, he expressed some criticisms especially focusing on the physical 
significance of considering the mutual information and the booklet 
construction in models with infinite range interactions.

We believe that we thoroughly addressed all the criticisms and suggestions of 
the referees in the new version of the manuscript.

Moreover, here below we provide a detailed answer to all the criticisms of the 
referees.

Sincerely,

Vincenzo Alba
Stephen Inglis 
Lode Pollet

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE FIRST REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------
	This paper deals with the study of entropies in classical spin models.
	The focus is on the Sherrington-Kirkpatrick model, a simple model that
	exhibits spin glass physics. Both analytical and numerical results are
	presented, and compared.

	Overall I found the paper interesting, and clearly written. Such
	classical entropies had only been studied in nearest neighbor spin
	models, so this is a useful addition to the existing literature.

RESPONSE:
========
	We thank the referee for considering our work interesting and 
	clearly written. We also would like to stress that classical entropies 
	have not been studied yet in disordered spin models, which is another 
	reason why we believe that our work is interesting.

REFEREE:
-------

	There is one issue that needs to be addressed I think. In part II the
	classical Renyi entropy is defined by Eq. (5), and this is what is
	studied in the rest of the paper. It is not clear to me which is the
	underlying probability distribution, i.e it is not clear to me that
	this equals Eq. (2), because of the long-range interactions. This
	point needs to be clarified, as using the term Renyi entropy would be
	a slight abuse of terminology, and the reader should be made aware of
	that.

RESPONSE:
========

	We apologize for not being very clear about the equivalence between (5) 
	and (2). Eq. (5) is equivalent to Eq. (2) even with the long range 
	interactions. This is now clarified in the first paragraph in sec II. 
	Notice that the only physical effect of long range interactions is that 
	there is no area law, in contrast with local models, as it is discussed 
	extensively in the manuscript.

REFEREE:
-------

	I also have a couple of minor remarks and suggestions, which I list
	below:

	1) In the long introduction, several values of \beta are mentioned. It
	might be helpful to recall the values of $\beta_c(w)$.

RESPONSE:
========

	In order to meet the suggestion of the second referee we moved the summary of 
	results to the conclusions. Therefore, in the new version of the manuscript 
	the referee suggestion is implemented in the conclusions. First, we should 
	clarify that $\beta$ always refers to the explicit temperature in a single 
	replica, and never to the effective temperature on the booklet. We thank 
	the referee for noticing that we do not mention  explicitly the values of 
	the critical temperature for the model on the booklet. We now mention explicitly 
	that the critical temperature is a decreasing function of the booklet aspect 
	ratio $\omega$ and that $\beta_c=1,1/2$ for $\omega=0,1$. Notice that  
	all the  values of beta that we mention in the summary of the results 
	(\beta=1,\beta=3) correspond to the low-temperature phase of the model on 
	the booklet.

REFEREE:
-------

	2) Maybe at the $\forall i \in A$ also in equation (8) ?

RESPONSE:
========

	We thank the referee for noticing that. We moved the $\forall i...$  
	in equation (8).

REFEREE:
-------

	3) Before (40), the sentence "..., one obtains a standard Landau
	theory" reads awkwardly.

RESPONSE:
========

	We removed the sentence in question.

REFEREE:
-------
	4) Could the authors comment somewhere on the Monte Carlo statistical
	errors when they are not shown?
	This is: 
	-Fig 8 (hidden behind some of the points, so you could say smaller than 
	point size in caption).
	-Fig 9 (same as 8)
	In the other figures the errors bars are fairly clear, but Fig 5 might be 
	non-obvious at first glance

RESPONSE:
========

	We thank the referee for the useful comment. In the captions of Fig 5, Fig 8, 
	and Fig 9 we now mention explicitly that the error bars are smaller than the 
	symbol size.

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
RESPONSE TO THE SECOND REFEREE 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

REFEREE:
-------

	In the manuscript BG13275, the authors study the Reny entropy and
	mutual information in the Sherrington-Kirkpatrick model. Using the
	replica method, the authors constructed the replica-symmetric solution
	and 1-step replica-symmetry-breaking solutions. They are solved
	numerically and are compared with the results of the Monte Carlo
	simulations. It was shown that the one-step RSB solution is enough to
	find a good agreement with the Monte Carlo result. The aspect-ratio
	and temperature dependences of the entropy and mutual information are
	analyzed in detail to see the scaling in terms of the system size and
	the critical properties.

RESPONSE:
========

	We thank the referee for the accurate summary of our results.

REFEREE:
-------

	The manuscript is written very clearly and the analytical calculations
	as well as numerical ones look reliable. The authors followed the
	standard procedure of the replica analysis and several ansatz such as
	(28) look reasonable.

RESPONSE:
========

	We thank the referee for finding our paper well written and our analysis 
	reliable.

REFEREE:
-------

	Although the analysis looks quite reasonable, the authors do not
	discuss the results thoroughly nor their physical meaning. Therefore,
	I can recommend the publication if the authors clarify following
	several questions.

	(1). What is the significance to treat infinite-range models in the
	problem of mutual information? As the authors mentioned in the
	manuscript, there is no notion of space and locality in infinite-range
	models. It is not clear for me to consider the booklet geometry in
	such systems. It would be helpful if the authors answer this simple
	question in the manuscript.

RESPONSE:
========

	We thank the referee for the opportunity to clarify this point. The 
	definition of the mutual information is based on formula (2) in the 
	manuscript. Formula (2) is the standard definition of the classical 
	Renyi entropy, which in the limit $n\to 1$ become the Shannon entropy. 
	Formula (2) and the Shannon entropy can be used in systems that do not 
	have the notion of space and locality. The two subsystems A and B 
	appearing in (2) can be interpreted as two complementary groups of 
	spins, without any need for an underlying spatial bipartition. Thus, 
	the quantities that we address in our study, e.g., the mutual information, 
	are well-grounded from an information theoretical point of view. This 
	is now stressed in the introduction below formula (2) with the sentence

		" For local spin models the bipartition correspond to a spatial 
		separation between the spins. However, the definition~\eqref{renyi-intro} 
		can be used  in models with no notion of space, where it quantifies 
		the correlation between two groups of spins rather than two spatial 
		regions."

	As for the booklet construction, (2) can be interpreted as ratios of partition 
	functions defined on the booklet. This is an exact correspondence, and it 
	is not related to any notion of locality. To clarify this point we now demonstrate 
	explicitly the equivalence between the booklet construction and (2) in a new paragraph 
	at the beginning of section 2. From the derivation it is clear that the booklet 
	construction is meaningful also in mean-field models.

	A further direct check is provided in appendix B where we discuss (following 
	the referee suggestion) the classical Renyi entropy and the mutual information 
	for the infinite-range clean (i.e., without disorder) Ising model. In Figure 12 
	and Figure 13 the continuous lines are obtained using the booklet 
	construction, while the symbols (rhombi) are the results using the definition 
	of the classical Renyi entropy and mutual information (formulas (2) and (4) 
	in the manuscript).

REFEREE:
-------

	Is there any other work treating the booklet geometry in mean field 
	systems? For example, what is obtained for the simple, pure Ising model 
	with infinite-range interaction?

RESPONSE:
========

	We would like to thank the referee for this very useful comment. The only study we 
	are aware of is in

		J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

	In this work the authors investigate the mutual information in the Lipkin-Meshkov-Glick 
	(LMG) model. In particular, they provide results for the case with zero magnetic field, 
	where the model becomes equivalent to a classical infinite-range Ising model. However, 
	the authors investigate the Shannon mutual information (i.e., the limit $n\to1$) 
	instead of the Renyi. This paper is now mentioned in the manuscript (Ref. [46]).

	Motivated by the referee observation and the results of Ref. [46] we now discuss 
	in the appendix B of the manuscript the Renyi entropy and mutual information in the 
	clean infinite-range Ising model. We find that, both exhibit a volume law for any 
	temperature. To compare with the SK model we now include in Figure 8 and 9 the results 
	in the thermodynamic limit for S_2/N and I_2/N for the clean model.

	Interestingly, in the clean model I_2 exhibits subleading (i.e., subextensive) 
	contributions. Specifically, these corrections are logarithmically divergent with 
	system size N and are present only for three values of the inverse temperature. Their 
	precise behaviors is given as

		beta=1/2     ---->  1/4 log(N)

		beta=1/(1+w) ---->  1/2 log(N)

		beta=1	     ---->  1/2 log(N)

	Notice that beta=1/(1+w) corresponds to the critical temperature of the model on the 
	2-sheets booklet with aspect ratio w.

	This is different for the Shannon mutual information (see Ref. [46]), which is finite at 
	any temperature except at the critical point $beta=1$ where it diverges logarithmically 
	with system size.

	These new results raise the question whether similar subleading corrections are 
	present in the S-K model, as also noticed by the referee. It would be 
	interesting to determine the prefactor of these logarithmic corrections to check 
	whether they are affected by the disorder. Additionally, for the SK model it is 
	interesting to understand whether the replica-symmetry-breaking mechanism affects the 
	prefactor of the subleading logarithmic corrections of the mutual information.

	However, in order to investigate analytically these issues one has to go beyond the saddle 
	point approximation, which, at least in the low-temperature region, requires knowing the full 
	replica-symmetry-broken solution of the model.

	Another possibility is to focus at beta=1/2 where a logarithmic correction is expected, 
	and where the high-temperature replica symmetric result that we provide fully captures 
	the saddle point behavior. The strategy would be to consider the deviations from the saddle 
	point in the Monte Carlo data. We checked that, indeed, our numerical results are suggestive 
	of a logarithmic subleading correction although further numerical effort would be needed 
	to clarify this point. We will certainly address this issue in future work.

REFEREE:
-------
	(2). The definition of the mutual information in the present case is
	not clear for me. Please describe how to define and calculate S(B) and
	S(A or B) in the SK model.

RESPONSE:
========

	We thank the referee for this useful comment. Eq. (2), and later Eq. (5) via the replica 
	trick, are both quite explicit. Combined with Eq. (4) they give the mutual information in 
	the SK model. The non-locality of the interactions does not affect the definition compared 
	with the usual one for local models. The definition of region A is however less obvious, 
	and in the case of this work, we simply take it as an arbitrary set of points, rather than 
	one that is spatially connected. As it is clear from the manuscript the best way to calculate 
	the mutual information is using (4) calculating the three different contributions S(A) S(B) 
	and S(AUB) separately. They correspond to ratios of partition functions on the 
	booklet with different omega as

		S_AUB = Z(omega=1)/Z(omega=0),

		S_A = Z(omega)/Z(omega=0)

		S_B = Z(1-\omega)/Z(omega=0)

	In the previous version of the manuscript it was not clearly stated how to obtain S_B and S_AUB. We 
	now have written it explicitly in section 2

		"The mutual information ${\mathcal I}_n(A,B,\beta)$ is obtained in terms of the booklet 
		partition functions using Eq.~\eqref{renyi} and Eq.~\eqref{MI}, where $S_n(B)$ is 
		obtained from~\eqref{renyi} by exchanging $A$ and $B$. Notice also that in~\eqref{renyi}  
		$Z(A\cup B,n,\beta)\equiv Z(n\beta)$ is the partition function of the model on the booklet 
		with all the $n$ sheets identified, equivalently on a single sheet but at temperature $n\beta$."

	We also added the sentence

		...Here we restrict ourselves to the second R\'enyi entropy $[S_2]$, which is obtained 
		in terms of the booklet partition functions $Z(\omega,2,\beta)$ (see II) as 
		$-\log(Z(\omega,2,\beta)/Z(0,2,\beta))$...

	at the beginning of section VII where we first discuss the Monte Carlo data for the classical 
	Renyi to stress its definition in terms of booklet partition function.

	From the numerical point of view, the strategy to evaluate the ratios of booklet partition 
	functions is outlined Appendix C.

REFEREE:
-------

	(3). The authors summarized that the entropy is proportional to the system size by using the 
	numerical analysis. Since the entropy is calculated from the logarithm of the partition function, 
	the result looks trivial for me. As far as I understand, the subleading term is	important to find 
	the boundary law behavior in the local (non mean-field) models. (The leading term is canceled out 
	by the definition of the mutual information). Do the authors consider that the subleading term is 
	not important in the present system? I expect that some universal properties exist in the subleading 
	terms.

RESPONSE:
========

	We agree that in general the leading term is proportional to the system size is an obvious result. 
	In the manuscript we report the entropy only as a sanity check of our calculations.  We also agree 
	that it is likely that the subleading terms contain more universal information. In fact, as we now 
	discuss in appendix B this is the case for the regular infinite-range Ising model. We thank the referee 
	for raising this point. However, although the analytical calculation of the subleading corrections to 
	the entropy and the mutual information is possible in principle, it requires going beyond the saddle point 
	approximation described in section III C. This should be easier in the paramagnetic (less interesting) 
	phase but it would require a considerable amount of work below the critical point. As we also mentioned 
	in our response to question (2) of the referee, we plan to address this issue in future work.

REFEREE:
-------

	(4). The authors study the aspect-ratio dependence of the result closely. It is clearly understood that 
	the limits omega=0 and 1 reduces to the standard results. This means that we can find a result interpolating 
	their limits in the present problem. I do not expect any interesting behavior between the limits. What 
	did the authors expect as the result and what do the authors consider the obtained results interesting, 
	other than the crossing behavior?

RESPONSE:
========

	We agree that omega=0 and 1 are trivial results, at least in the standard SK model, and that one would 
	expect that the results between these two points should vary smoothly, and this is reflected in our data. 
	However, different aspect ratios omega are needed in the booklet constructions in order to construct the 
	mutual information, which is the main reason why we present them. We should also stress that the functional 
	dependence of the mutual information on omega is not trivial. The same statement of the referee would apply 
	to local models. However, as it is clear from Ref [29], the shape function of the mutual information 
	contains universal behavior if the parent model is critical and described by a conformal field theory. 
	In that respect, we believe it was worth examining the aspect ratio behavior in the SK model.

REFEREE:
-------

	(5). This comment is related to the previous one (4). Strictly
	speaking, I do not see any surprising behavior in the result. What is
	the significance of the present study? Did the authors find anything
	on the information properties of the SK model?

RESPONSE:
========

	From our answer to point 1) it should be clear now that our work is well grounded from the information theoretical 
	point of view as we are studying the standard classical Renyi entropies and mutual information. These 
	quantities have been shown to contain universal information for instance in local critical models described 
	a conformal field theory (Ref. [29]). Again, we should stress that their behavior in disordered and glassy 
	systems has not been investigated in the previous literature. We chose to focus on the SK model which is one of 
	the few exactly solved models for spin glasses and the base case for understanding glassy behavior. Also its 
	solution exhibits quite complex features (for instance, replica symmetry breaking). We find already quite 
	interesting that these complex features are reflected in the leading behavior of the mutual information and 
	entropy.

	Moreover, our study can be considered a zeroth order starting point because, as correctly observed by the 
	referee and as it is clear from our results on the clean infinite range Ising model (appendix B), further 
	interesting information could be contained in the finite-size effects of the mutual information. On the other 
	hand our study characterizing the extensive part of the mutual information provides the necessary background 
	for conducting these further studies.

REFEREE:
-------

	(6). The authors only focused on the result at n=2. It would be	interesting if the authors mention the n-dependence of 
	the result. Especially, the limit n=1 would give a nontrivial result and is interesting. I expect that it wouldn't 
	be difficult to analyze the replica symmetric solution.

RESPONSE:
========

	We thank the referee for asking about this point. In fact we thought about investigating the analytic continuation 
	n->1 to obtain the Shannon Renyi entropy and mutual information. This question is interesting especially in order to 
	compare with the results for the clean infinite range Ising model obtained in Ref. [46],  where it has been shown 
	that the Shannon mutual information is O(1) (i.e., not extensive) at any temperature except for the critical point at 
	beta=1 where it diverges logarithmically with system size.

	Following the referee suggestion, we now provide results for the analytic continuation n->1. In the new version of 
	the manuscript in section IV B we discuss the analytic continuation for the paramagnetic replica symmetric case, 
	although this could be done as well in the replica symmetry broken region of the phase diagram.

	From the solution we observe that in the limit of infinite temperature one has

	   I_1/N = beta^2 omega(1-omega)/2

	i.e., the Shannon mutual information obeys a volume law. This is in contrast with the result in the clean Ising 
	reported in Ref. [46]. This suggests that the subextensive behavior for I_1 found for the clean system is an effect 
	of fine tuned cancellations that do not happen in presence of disorder.

REFEREE:
-------

	(7). Although the manuscript is written very clearly, it is somewhat
	too long. For example, the authors summarized the results in the
	introduction, and they are described again in the final section. It is
	not necessary to summarize the results in the introduction.

RESPONSE:
========

	We would like to thank the referee for this useful comment. We now summarize the main results of the paper only in the 
	conclusions.

REFEREE:
-------
	(8). J_{ij} should be inserted in Eq.(10).

RESPONSE:
========

	We thank the referee for spotting this. We restored the missing J_{ij} in 
	Eq. (10).

REFEREE:
-------

	(9). I am not sure whether this manuscript is suitable to PRB. I would think that PRE is more appropriate.

RESPONSE:
========

	We believe that after our detailed answer to the referee points it is clear that our work addresses the issue of how 
	standard information-theory motivated quantities such as entropies and mutual information are affected by disorder and 
	glassy criticality. Disordered magnetic systems and glassy phase transitions are traditionally topics of PRB, as 
	well as (especially in recenty years) entanglement and phase transitions. Therefore we believe that our paper 
	is appropriate for PRB.

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
SUMMARY OF MAIN CHANGES IN THE MANUSCRIPT 
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

1) Shortened introduction. Now the summary of our main results is in the Conclusions, 
   as suggested by the second referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

2) Modified the captions of Figure 5,8,9 to meet the suggestions of the first referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

3) Added the results for the entropy and mutual information in the infinite-range 
   clean Ising model in Figure 8,9.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

4) Added appendix B (with 5 Figures) to discuss the entropy and mutual information 
   in the clean infinite-range Ising model to meet the suggestions of the second 
   referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

5) Added subsection VI B to discuss the analytic continuation $n\to 1$, as suggested 
   by the second referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

6) Added reference [46]

	J. Wilms, J. Vidal, F. Verstraete, S. Dusuel, J. Stat. Mech. P01023 (2012)

   to previous work on the Shannon mutual information in the infinite-range Ising model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

7) Added

	$\forall i\in A$

   in Eq (8) as suggested by the first referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

8) Removed the sentence

	"..., one obtains a standard Landau theory"

   before Eq (40) in the previous version of the manuscript, as suggested by the first 
   referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

9) Added the paragraph below formula (3)

	" For local spin models the bipartition correspond to a spatial 
	separation between the spins. However, the definition~\eqref{renyi-intro} 
	can be used  in models with no notion of space, where it quantifies 
	the correlation between two groups of spins rather than two spatial 
	regions."

   to meet the suggestion of the second referee.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

10) Added the paragraph

	"...Given a generic \emph{classical} spin model at inverse temperature $\beta$, 
	The probability $p_i$ of a given spin configuration $i\in{\mathcal C}$, with ${\mathcal 
	C}$ being the set of all possible configurations, is given by the Boltzmann weight 
	$p_i=e^{-\beta E(i)}/Z$, with $E(i)$ the associated energy. In presence of a 
	bipartition (see Fig.~\ref{cartoon} (a)) the probability of a spin configuration 
	$i_A\in{\mathcal C}_A$ is given as 
	$p_{i_A}=\sum_{i_B}e^{-\beta E(i_A,i_B)}/Z$, where the sum is over all possible 
	spin configurations in part $B$. Note that this is valid for generic interactions, 
	i.e., both local and non-local ones. Clearly, one has $p^n_{i_A}\equiv\frac{1}{Z^n}
	\sum_{(i^1_B,\dots,i^n_{B})}e^{-\beta\sum_k E(i_A,i^k_{B})}$...."

   at the beginning of section II to meet the suggestions of both the referees.